---
title: "TwoSided T-Test and Boxplots of Final Findings"
author: "Rosa G., Lilian G., Sylvia N."
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### OR Statistical Analysis (5 and 6 mutations, with mutation A)

```{r,include=FALSE}
# Hemoglobin of 5 mutations with mutation A
#5 [((2, 3, 4, 12), 198.0), ((3, 4, 6, 7), 176.0), ((4, 5, 7, 9), 164.0), ((2, 3, 5, 6), #149.0), ((2, 3, 5, 6), 155.0)]

#Hemoglobin of 6 mutations with mutation A

#6 [((2, 3, 5, 6, 12), 208.0), ((2, 3, 4, 5, 7), 205.0), ((3, 5, 6, 7, 9), 148.0), ((2, 3, 4, 5, 6), 171.0), ((2, 3, 4, 5, 6), 152.0), ((2, 3, 4, 6, 7), 176.0), ((2, 3, 5, 7, 9), 147.0), ((2, 6, 7, 9, 12), 154.0), ((2, 3, 4, 5, 6), 145.0), ((2, 4, 5, 6, 7), 189.0), ((2, 3, 4, 5, 11), 165.0), ((2, 3, 4, 5, 7), 113.0), ((2, 3, 6, 7, 12), 153.0), ((2, 3, 5, 7, 11), 124.0), ((2, 3, 5, 6, 7), 148.0)]


```


```{r,include=FALSE}
# hemoglobin levels (5 and 6 mutations)
five_mut_or=c(198,176,164,149,155) #len=5
six_mut_or=c(208,205,148,171,152,176,147,154,145,189,165,113,153,124,148) #;en=15

# hemoglobin levels (all 5 and 6 mutations combined)
or_hemoglobin=c(five_mut_or,six_mut_or) #len=20
```

#### First finding: mutation M has positive contribution to hemoblobin

```{r,include=FALSE}
#Hemoglobin of 5 mutations with mutation A
#5 [((2, 3, 4, 12), 198.0), ((3, 4, 6, 7), 176.0), ((4, 5, 7, 9), 164.0), ((2, 3, 5, 6), 149.0), ((2, 3, 5, 6), 155.0)]

#Hemoglobin of 6 mutations with mutation A

#6 [((2, 3, 5, 6, 12), 208.0), ((2, 3, 4, 5, 7), 205.0), ((3, 5, 6, 7, 9), 148.0), ((2, 3, 4, 5, 6), 171.0), ((2, 3, 4, 5, 6), 152.0), ((2, 3, 4, 6, 7), 176.0), ((2, 3, 5, 7, 9), 147.0), ((2, 6, 7, 9, 12), 154.0), ((2, 3, 4, 5, 6), 145.0), ((2, 4, 5, 6, 7), 189.0), ((2, 3, 4, 5, 11), 165.0), ((2, 3, 4, 5, 7), 113.0), ((2, 3, 6, 7, 12), 153.0), ((2, 3, 5, 7, 11), 124.0), ((2, 3, 5, 6, 7), 148.0)]
```



```{r, echo=FALSE}
five_mut_or_mutM=c(198)
six_mut_or_mutM=c(208,154,153)
or_hemoglobin_mutM=c(five_mut_or_mutM,six_mut_or_mutM)

## two sided t test
# compute t-test for diff. of means
t.test(or_hemoglobin,or_hemoglobin_mutM,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(or_hemoglobin,or_hemoglobin_mutM,main="Hemoglobin Levels (OR), 5 and 6 mutations",names=c("All Individuals with mutation A","Individuals with AM Coalition"),ylab="Hemoglobin Levels")
```


#### Second finding: Coalition HM have a negative contribution to hemoglobin

```{r,include=FALSE}
#Hemoglobin of 5 mutations with mutation A
#5 [((2, 3, 4, 12), 198.0), ((3, 4, 6, 7), 176.0), ((4, 5, 7, 9), 164.0), ((2, 3, 5, 6), 149.0), ((2, 3, 5, 6), 155.0)]

#Hemoglobin of 6 mutations with mutation A

#6 [((2, 3, 5, 6, 12), 208.0), ((2, 3, 4, 5, 7), 205.0), ((3, 5, 6, 7, 9), 148.0), ((2, 3, 4, 5, 6), 171.0), ((2, 3, 4, 5, 6), 152.0), ((2, 3, 4, 6, 7), 176.0), ((2, 3, 5, 7, 9), 147.0), ((2, 6, 7, 9, 12), 154.0), ((2, 3, 4, 5, 6), 145.0), ((2, 4, 5, 6, 7), 189.0), ((2, 3, 4, 5, 11), 165.0), ((2, 3, 4, 5, 7), 113.0), ((2, 3, 6, 7, 12), 153.0), ((2, 3, 5, 7, 11), 124.0), ((2, 3, 5, 6, 7), 148.0)]
```

```{r,echo=FALSE}
five_mut_or_mutHM=c()
six_mut_or_mutHM=c(154,153)
or_hemoglobin_mutHM=c(five_mut_or_mutHM,six_mut_or_mutHM)

## two sided t test
# compute t-test for diff. of means
t.test(or_hemoglobin,or_hemoglobin_mutHM,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(or_hemoglobin,or_hemoglobin_mutHM,main="Hemoglobin Levels (OR), 5 and 6 mutations",names=c("All Individuals with mutation A","Individuals with AHM Coalition"),ylab="Hemoglobin Levels")



```

### AND Statistical Analysis (2-6 muations)

```{r,include=FALSE}
## each list looks like: [[1]] [[1]][[1]][1] 2, [[1]][[2]][1] 9, [[1]][[3]][1] 195

two_mutations=list( c(list(2, 9), 195.0), c(list(2,5), 185.0), c(list(0, 6), 156.0), c(list(0, 4), 78.0), c(list(2, 7), 79.0), c(list(0, 9), 199.0), c(list(3, 6), 167.0), c(list(0, 6), 145.0), c(list(1, 7), 179.0), c(list(0, 6), 199.0), c(list(0, 4), 113.0), c(list(0, 7), 187.0), c(list(2, 3), 170.0), c(list(2, 7), 124.0), c(list(0, 2), 148.0), c(list(0, 4), 147.0), c(list(0, 6), 155.0))
 

three_mutations= list(c(list(1, 2, 5), 248.0), c(list(2, 3, 7), 193.0), c(list(0, 2, 3), 171.0), c(list(0, 2, 3), 168.0), c(list(1, 2, 3), 198.0), c(list(0, 2, 5), 159.0), c(list(1, 5, 7), 143.0), c(list(7, 9, 12), 153.0), c(list(0, 3, 7), 173.0), c(list(2, 4, 6), 172.0), c(list(3, 7, 9), 138.0), c(list(3, 6, 9), 168.0), c(list(2, 7, 12), 129.0), c(list(0, 2, 3), 152.0), c(list(0, 3, 9), 148.0), c(list(0, 4, 6), 176.0), c(list(0, 4, 7), 164.0), c(list(0, 3, 5), 111.0), c(list(0, 2, 7), 147.0), c(list(3, 4, 7), 144.0), c(list(0, 4, 5), 152.0), c(list(0, 6, 7), 141.0), c(list(0, 2, 3), 149.0), c(list(2, 7, 9), 129.0), c(list(0, 2, 3), 165.0), c(list(2, 3, 6), 153.0), c(list(2, 3, 12), 146.0) )
 

four_mutations= list(c(list(0, 2, 3, 6), 196.0), c(list(0, 3, 7, 11), 149.0), c(list(0, 2, 3, 9), 189.0), c(list(0, 3, 7, 12), 231.0), c(list(0, 4, 6, 7), 241.0), c(list(0, 2, 3, 6), 199.0), c(list(2, 3, 5, 7), 205.0), c(list(2, 3, 7, 12), 119.0), c(list(0, 4, 6, 7), 128.0), c(list(0, 4, 5, 12), 140.0), c(list(0, 2, 3, 4), 155.0), c(list(2, 3, 4, 6), 170.0), c(list(0, 2, 3, 12), 63.0), c(list(0, 2, 5, 11), 171.0), c(list(0, 2, 3, 7), 142.0), c(list(0, 2, 3, 7), 151.0), c(list(2, 4, 6, 9), 143.0), c(list(0, 3, 4, 5), 171.0), c(list(0, 2, 3, 4), 152.0), c(list(0, 2, 6, 12), 167.0), c(list(0, 2, 3, 11), 149.0), c(list(2, 3, 6, 7), 176.0), c(list(0, 2, 3, 4), 181.0), c(list(0, 2, 6, 7), 154.0), c(list(2, 3, 5, 6), 162.0), c(list(2, 5, 6, 9), 130.0), c(list(2, 3, 7, 9), 164.0), c(list(0, 2, 4, 9), 190.0), c(list(0, 2, 3, 4), 165.0), c(list(0, 3, 4, 9), 140.0), c(list(2, 4, 7, 9), 159.0), c(list(2, 3, 6, 7), 164.0), c(list(0, 2, 3, 12), 114.0), c(list(0, 2, 3, 9), 104.0), c(list(0, 4, 7, 9), 163.0), c(list(0, 2, 3, 7), 133.0), c(list(0, 2, 7, 9), 209.0), c(list(0, 2, 3, 7), 145.0))
 

five_mutations= list(c(list(0, 2, 3, 4, 7), 181.0), c(list(0, 2, 3, 4, 9), 181.0), c(list(0, 2, 4, 6, 7), 169.0), c(list(0, 3, 5, 6, 12), 208.0), c(list(0, 2, 3, 6, 9), 203.0), c(list(0, 2, 3, 4, 7), 185.0), c(list(0, 3, 5, 6, 12), 201.0), c(list(0, 2, 4, 6, 12), 158.0), c(list(0, 3, 4, 5, 7), 185.0), c(list(0, 3, 5, 9, 12), 180.0), c(list(0, 2, 3, 6, 9), 186.0), c(list(2, 5, 7, 9, 12), 166.0), c(list(0, 3, 4, 5, 6), 150.0), c(list(0, 4, 6, 7, 9), 90.0), c(list(2, 3, 4, 6, 7), 172.0), c(list(2, 3, 5, 7, 11), 159.0), c(list(3, 4, 6, 7, 12), 167.0), c(list(2, 3, 5, 6, 7), 134.0), c(list(0, 3, 5, 6, 7), 148.0), c(list(2, 3, 4, 6, 7), 175.0), c(list(0, 2, 3, 4, 12), 134.0), c(list(0, 2, 3, 7, 9), 144.0), c(list(0, 1, 4, 9, 12), 149.0), c(list(2, 3, 5, 9, 12), 169.0), c(list(2, 3, 4, 6, 7), 192.0), c(list(0, 2, 3, 5, 12), 154.0), c(list(0, 4, 9, 11, 12), 184.0), c(list(0, 2, 3, 5, 7), 138.0), c(list(0, 3, 6, 7, 12), 174.0), c(list(0, 1, 2, 3, 7), 187.0), c(list(0, 4, 6, 7, 12), 157.0), c(list(0, 3, 4, 5, 6), 164.0), c(list(0, 3, 4, 6, 9), 134.0), c(list(0, 6, 7, 9, 12), 130.0), c(list(0, 2, 5, 6, 7), 168.0)) 
 

six_muations=list(c(list(0, 2, 3, 5, 7, 9), 165.0), c(list(0, 2, 4, 5, 7, 12), 165.0), c(list(0, 2, 3, 5, 7, 9), 169.0), c(list(0, 2, 3, 5, 6, 9), 149.0), c(list(0, 1, 2, 4, 6, 7), 175.0), c(list(0, 2, 4, 5, 6, 7), 189.0), c(list(0, 2, 3, 4, 5, 9), 138.0), c(list(2, 3, 4, 6, 7, 9), 150.0), c(list(0, 2, 3, 7, 8, 9), 130.0), c(list(0, 1, 2, 3, 7, 9), 165.0), c(list(0, 2, 3, 6, 11, 12), 140.0))

and_mutations=c(two_mutations,three_mutations,four_mutations,
                five_mutations,six_muations)


```



```{r,include=FALSE}
########### FUNCTION TO GET VECTOR OF ALL HEMOGLOBIN LEVELS  #######

all_mutation_hemolevels<-function()
{
  hemoglobin_levels=c()
  ## will go from 1-128 (all 2-6 mutation groupings)
  #### Ex. [[128]] is the last part in list,
  #### so it should be of 6 mutations with hemo. level
  #### so length is 7, and that is where the hemo. level is and we want to save
  #### 
  for(i in 1:length(and_mutations))
  {
    ## puts hemoglobin level in vector
    hemoglobin_levels<-c(hemoglobin_levels,
                         and_mutations[[i]][[length(and_mutations[[i]])]])
  }
  return(hemoglobin_levels)
}
```




```{r,include=FALSE}
########### FUNCTION TO GET VECTOR OF HEMOGLOBIN LEVELS BASED ON COALITION #######

coalition_mutation_hemolevels<-function(spec_coalition)
{
  hemoglobin_levels=c()
  ## will go from 1-128 (all 2-6 mutation groupings)
  #### Ex. [[128]] is the last part in list, so it should be of 6 mutations with hemo. level
  #### so length is 7, but we want to look at 1-6 (mutation vals.), 7 is hemo.
  #### 
  for(i in 1:length(and_mutations))
  {
    ### will go from 1st mutation val to last mutation val
    ## will be empty each time we go through the whole list
    mutations=c()
   for(j in 1:(length(and_mutations[[i]])-1))
   {
     ## puts all mutations in vector
     mutations=c(mutations,and_mutations[[i]][[j]])
   }
    ## checks if coalition is in the vector (all must be there to get hemo. level)
    if(all(spec_coalition %in% mutations)==TRUE)
      {
      hemoglobin_levels<-c(hemoglobin_levels,and_mutations[[i]][[length(and_mutations[[i]])]])
    }
    
  }
  return(hemoglobin_levels)
}

```


```{r,include=FALSE}
## Vector of all hemoglobin levels from groupings of 2-6 mutations (128 people)
AND_all_hemoglobin_levels<-all_mutation_hemolevels()
```


#### Third finding: Coalition DG has a positive contribution to hemoglobin levels

```{r,echo=FALSE}
#D<-3, G<-6
DG_hemo_levels=coalition_mutation_hemolevels(c(3,6))

## two sided t test
# compute t-test for diff. of means
t.test(AND_all_hemoglobin_levels,DG_hemo_levels,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(AND_all_hemoglobin_levels,DG_hemo_levels,main="Hemoglobin Levels (AND), 2-6 mutations",names=c("All Individuals","Individuals with DG Coalition"),ylab="Hemoglobin Levels")

```


#### Fourth finding: Coalition CM has a negative contribution to hemoglobin levels

```{r,echo=FALSE}
#C<-2, M<-12
CM_hemo_levels=coalition_mutation_hemolevels(c(2,12))

## two sided t test
# compute t-test for diff. of means
t.test(AND_all_hemoglobin_levels,CM_hemo_levels,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(AND_all_hemoglobin_levels,CM_hemo_levels,main="Hemoglobin Levels (AND), 2-6 mutations",names=c("All Individuals","Individuals with CM Coalition"),ylab="Hemoglobin Levels")


```

#### Fifth finding: Coalition BCD has a positive contribution to hemoglobin levels

```{r,echo=FALSE}
#B<-1, C<-2, D<-3
BCD_hemo_levels=coalition_mutation_hemolevels(c(1,2,3))

## two sided t test
# compute t-test for diff. of means
t.test(AND_all_hemoglobin_levels,BCD_hemo_levels,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(AND_all_hemoglobin_levels,BCD_hemo_levels,main="Hemoglobin Levels (AND), 2-6 mutations",names=c("All Individuals","Individuals with BCD Coalition"),ylab="Hemoglobin Levels")



```



#### Sixth finding: Coalition ACDM has a negative contribution to hemoglobin levels

```{r,echo=FALSE}
#A<-0 , C<-2 . D<-3 , M<-12
ACDM_hemo_levels=coalition_mutation_hemolevels(c(0,2,3,12))

## two sided t test
# compute t-test for diff. of means
t.test(AND_all_hemoglobin_levels,ACDM_hemo_levels,alternative = c("two.sided"))
# Note: this leads you to reject the null hypothesis that the means are the same.
# make a primitive boxplot
boxplot(AND_all_hemoglobin_levels,ACDM_hemo_levels,main="Hemoglobin Levels (AND), 2-6 mutations",names=c("All Individuals","Individuals with ACDM Coalition"),ylab="Hemoglobin Levels")




```




